---
title: "Analysis Report 1: NEON Small Mammal Data"
author: "Alexia Ford"
date: "3/15/2018"
bibliography: references.bib
csl: ecology.csl
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This report is testing how the distrubition of weight varies across NEON sites using the National Ecological Observatory Network (NEON) data. This report will explain NEON's work/purpose, NEON's design, introduce the data used, test a hypothesis, and analyze the information presented. The National Ecological Observatory Network (NEON) is an ecological observation platform for discovering, understanding, and forecasting the impacts of climate change, land use change, and invasive species on continental-scale ecology with sites distributed within 20 eco-climatic domains across United States, Alaska, Hawaii, and Puerto Rico using remote sensing [@kampe2010neon]. NEON has allowed for easy accessible data to the public by providing the data the observatory has collected to scientists, land managers, policy makers, educators, and members of the general public interested in addressing these major ecological challenges. The data produced by NEON are strategically aimed at addressing ecological questions. Questions that are analyzing complex ecological proccesses and relationships that govern large scale patterns consistent with the measuring the drivers of change, biogeochemical cycling, and population and community responses across a variety of ecosystems in the ultimate hopes of expanding our understanding of ecological processes and to improve our ability to forecast these ecological changes [@thorpe2016introduction].

The project is responsible for measuring the causes and effects of environmental change across the United States within these 20 eco-climate domains that all represent different landforms, vegetation, climate, and ecosystem dynamics. The project is constantly collecting data from it’s observatory’s regional domain to provide data and statistically represent ecological, physical, and biological variability across the US. NEON has four different sites that the observatory collects data from, these sites include: core terrestrial sites, core aquatic sites, relocatable terrestrial sites, and lastly relocatable aquatic sites. The sites are collected  using five measurement subsystems of the Observatory including: the Airborne Observation Platform (AOP), Aquatic Instrument System (AIS), Aquatic Observation System (AOS), Terrestrial Instrument System (TIS), and Terrestrial Observation System (TOS) [@thorpe2016introduction]. These five systems of measurement have been used to collect endless data and NEON has continued to do so by properly standardizing protocols across all sites and maintaining community best practices. 
This report specifically will be looking at NEON's collection of small mammal data. NEON collects small mammal data by carefully following many protocols and sampling methods. The collection of small mammal field studies has played a vital role in ecology particulartly when studying behavioral, population, and community ecology. The use of small mammals is especially helpful as they are easily handled due to trapability and size, as well as readily abundant as they are present in various ecosystems. Small mammals are exceptionally suitable as they are canidate for both short and long-term studies. Short term studies such as examining resource limitation and preditation in population dynamics. Oppsingly, they are also canidates for long-term studies such as small mammal communities responses to extreme climatic events or cycles. 

This report will specifically be analyzing the weight of small mammals across all NEON sites, to then take a closer look at the site with the highest average weight and the site with the lowest average weight. The LBJ National Grassland ("CLBJ" site) located in Texas is the site with the highest average weight among small mammals. The Smithsonian Environmental Research Center ("SERC" site) located in Maryland is opposingly the site with the lowest average weight among small mammals. 

The LBJ National Grassland is a land mass consisting of 16,800 acres of land and is a NEON core terrestrial site located in the Southern Plains of Texas. The site has a rich history used for recreation and hunting, livestock grazing, and fossil fuel extraction. Ecological monitoring is continually performed at the site along with prescribed burning. The climate of this glassland is considered humid subtropical, and the weather typically consists of frequent variations both daily and seasonaly besides in the hot and humid summers. The average precipitation is 36.8 inches with average summer highs of 94F and average winter lows of 27F. The site has varying dominant tree and grass species, with a particularly clay-rich soil. Lastly, this grasslands wildlife species includes: pronghorn, black-tailed prarie dogs, horned larks, and mountain plovers. 

The Smithsonian Environmental Research Center is a relocatable terrestrial NEON site located in Maryland around the Mid-Atlantic coast. This site is surrounded by varying land types including oceanic, grasslands, forests, etc. This site is highly influenced by weather changes dictated greatly by the Atlantic ocean and greatlakes. The Smithsonian Environmental Research Center is an important terrestrial ecosystem that include shrublands that provide crucial habitat for wildlife, and communities such as limestone and dolomite glades that are home to endangered plant species [@rogers2000climate]. 

Delete this placeholder text, and replace it with your introduction text. You should include background on the NEON project including proper citations (e.g. [@thorpe2016introduction]), the sets of sites you are comparing, and other relevent background information related to your specific hypotheses. 

# Methods
The data for this report is collected by the National Ecological Observatory Network (NEON). NEON collects small mammal data by carefully following many protocols and sampling methods. NEON's typical target species are nocturnal and cryptic and the common method to sample them include: passigve, involving kil or live traps, and baited or unbaited. For long-term studies such as small mammal ccommunity, population, and disease dynamics mark-rapture methods are commonly used. The mark-rapture design includes: trap type, bait, marking method, and trap array. 

The trap type consists of box traps which are live traps that include: Sherman traps, Longworth traps, and Elliott traps that are placed on the ground or in trees and baited. The diversity of traps is necessary considering no trap type or size is appropriate for all species, for exmaple Sherman traps have a padded compartment that is more adept to cold conditions. 

Bait is an additional mark-recapture design that greatly increases small mammal capture success. The bait used in the mark-recapture method 

Since the target species for NEON are nocturnal and cryptic, the most common methods used to sample them are passive, involving kill- or live-traps, baited or unbaited (e.g., Wilson et al. 1996, Evangelista et al. 2008). For long-term studies of small mammal community, population, and disease dynamics, mark-recapture methods are the most commonly deployed
A number of analytical methods have been developed to use mark-recapture data to derive estimates of density and diversity, given that many animal species, including small mammals, cannot be observed perfectly 


Describe where the data came from, including how they were collected in the field (information is in the pdf file in the data folder), some information about those sites and why you chose them, and how you are analyzing them (e.g. using R to visualize the results and the statistical tests you will use).

# Results
```{r load-packages}
library("dplyr")
library("ggplot2")
library("readr")
```

Read in the small mammal data set:
```{r load-in-data}
# read in the small mammal data from raw data directory
mam_pertrapnight <- read_csv("data/raw_data/mam_pertrapnight.csv")
```

Box plot for the weight of small mammals at all NEON sites:
```{r boxplot-of-small-mammal-weight}
# weight of small mammals at diffrent sites in the NEON data
ggplot(mam_pertrapnight, aes(x = siteID,
                             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Small Mammals at all NEON sites") +
  xlab("Site ID") +
  ylab("Weight (grams)")
```

Box plot for the weight of small  mammals at "CLBJ" NEON site:
```{r boxplot-for-CLBJ-site}
# make a boxplot for the weight of small mammals at "CLBJ" NEON site
mam_pertrapnight %>%
  filter(siteID == "CLBJ") %>%
  filter(!is.na(scientificName)) %>%
  ggplot(aes(x = scientificName,
             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Small Mammal Species at CLBJ NEON site") +
  xlab("Species") +
  ylab("Weight (grams)")
```


Boxplot for the weight of small  mammals at "SERC" NEON site:
```{r boxplot-for-SERC-site}
# make a boxplot for the weight of small mammals at "SERC" NEON site
mam_pertrapnight %>%
  filter(siteID == "SERC") %>%
  filter(!is.na(scientificName)) %>%
  ggplot(aes(x = scientificName,
             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Small Mammal Species at SERC NEON site") +
  xlab("Species") +
  ylab("Weight (grams)")
```


```{r boxplot-of-neotoma-weight-all-sites}
mam_pertrapnight %>%
  filter(scientificName == "Neotoma micropus") %>%
  ggplot(aes(x = siteID,
             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Neotoma micropus Species at all NEON Sites") +
  xlab("Site") +
  ylab("Weight (grams)")
```

```{r boxplot-of-mus-weight-all-sites}
mam_pertrapnight %>%
  filter(scientificName == "Mus musculus") %>%
  ggplot(aes(x = siteID,
             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Mus musculus Species at all NEON Sites") +
  xlab("Site") +
  ylab("Weight (grams)")
```

Add a number of code chunks in the Results section. These should read in, subset and plot the data as needed (no need to save any figures to pdf, since they will be put into the rendered document when you click 'knit'), and, for any hypotheses that you want to test, an appropriate statistical test. You should have at least 5 figures and/or tables in this section, which you use to test your hypotheses against the dataset.

# Discussion

Discuss the results you found in the context of what you know about the site/sites and the predominant types of vegatation in those locations. Think also about the community dynamics of the mammal community at each of the sites, in the context of what you know about different types of species interactions, food webs, trophic transfer, etc.

# Sources Cited

