---
title: "Analysis Report 1: NEON Small Mammal Data"
author: "Alexia Ford"
date: "3/15/2018"
bibliography: references.bib
csl: ecology.csl
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This report is testing how the distrubition of weight varies across the National Ecological Observatory Network (NEON) sites using data collected and publically presented by NEON. This report will explain NEON's work/purpose, NEON's design, present the data used, test a hypothesis, and lastly analyze the results. The National Ecological Observatory Network is an ecological observation platform for discovering, understanding, and forecasting the impacts of climate change, land use change, and invasive species on continental-scale ecology with sites distributed within 20 eco-climatic domains across United States, Alaska, Hawaii, and Puerto Rico using remote sensing [@kampe2010neon]. NEON has allowed for easy accessible data to the public by providing the data the observatory has collected to scientists, land managers, policy makers, educators, and members of the general public interested in addressing these major ecological challenges. The data produced by NEON are strategically aimed at addressing ecological questions, questions that are analyzing complex ecological proccesses and relationships that govern large scale patterns consistent with the measuring the drivers of change, biogeochemical cycling, and population and community responses across a variety of ecosystems in the ultimate hopes of expanding our understanding of ecological processes and to improve our ability to forecast these ecological changes [@thorpe2016introduction].

The project is responsible for measuring the causes and effects of environmental change across the United States within these 20 eco-climate domains that all represent different landforms, vegetation, climate, and ecosystem dynamics. The project is constantly collecting data to provide and statistically represent ecological, physical, and biological variability across the US. NEON has four different sites that the observatory collects data from, these sites include: core terrestrial sites, core aquatic sites, relocatable terrestrial sites, and lastly relocatable aquatic sites. The sites are collected  using five measurement subsystems of the Observatory including: the Airborne Observation Platform (AOP), Aquatic Instrument System (AIS), Aquatic Observation System (AOS), Terrestrial Instrument System (TIS), and Terrestrial Observation System (TOS) [@thorpe2016introduction]. These five systems of measurement have been used to collect endless data and NEON has continued to do so by properly standardizing protocols across all sites and maintaining community best practices. 

This report will be looking at NEON's collection of small mammal data. NEON collects small mammal data by carefully following many protocols and sampling methods. The collection of small mammal field studies has played a vital role in ecology particularly when studying behavioral, population, and community ecology. The use of small mammals is especially helpful as they are easily handled due to trapability and size, as well as readily abundant as they are present in various ecosystems. Small mammals are exceptionally suitable as they are canidate for both short and long-term studies. Short term studies such as examining resource limitation and preditation in population dynamics. Oppsingly, they are also canidates for long-term studies such as small mammal communities responses to extreme climatic events or cycles. 

This report will specifically be analyzing the weight of small mammals across all NEON sites, to then take a closer look at the site with the highest average weight and the site with the lowest average weight. The LBJ National Grassland ("CLBJ" site) located in Texas is the site with the highest average weight among small mammals. The Smithsonian Environmental Research Center ("SERC" site) located in Maryland is, opposingly, the site with the lowest average weight among small mammals. 

The LBJ National Grassland is a land mass consisting of 16,800 acres of land and is a NEON core terrestrial site located in the Southern Plains of Texas. The site has a rich history used for recreation and hunting, livestock grazing, and fossil fuel extraction. Ecological monitoring is continually performed at the site along with prescribed burning. The climate of this glassland is considered humid subtropical, and the weather typically consists of frequent variations both daily and seasonaly besides in the constant hot and humid summers. The average precipitation is 36.8 inches with average summer highs of 94F and average winter lows of 27F. The site has varying dominant tree and grass species, with a particularly clay-rich soil. Lastly, this grasslands wildlife species includes: pronghorn, black-tailed prarie dogs, horned larks, and mountain plovers. 

The Smithsonian Environmental Research Center is a relocatable terrestrial NEON site located in Maryland around the Mid-Atlantic coast. This site is surrounded by varying land types including oceanic, grasslands, forests, etc. This site is highly influenced by weather changes dictated greatly by the Atlantic ocean and greatlakes. The Smithsonian Environmental Research Center is an important terrestrial ecosystem that includes shrublands that provide crucial habitat for wildlife, and communities such as limestone and dolomite glades that are home to endangered plant species [@rogers2000climate]. 

Delete this placeholder text, and replace it with your introduction text. You should include background on the NEON project including proper citations (e.g. [@thorpe2016introduction]), the sets of sites you are comparing, and other relevent background information related to your specific hypotheses. 

# Methods
The data for this report is collected by the National Ecological Observatory Network (NEON). NEON collects small mammal data by carefully following many protocols and sampling methods. NEON's typical target species are nocturnal and cryptic. The common method to sample these target species include: passigve, involving kill or live traps, and baited or unbaited. For long-term studies such as small mammal ccommunity, population, and disease dynamics mark-rapture methods are commonly used. NEON follows parts of the mark-rapture designs which includes: trap type, bait, marking method, and trap array.

The trap type NEON uses consists of box traps which are live traps that include: Sherman traps, Longworth traps, and Elliott traps that are placed on the ground or in trees and baited. NEON uses H.B. Sherman folding traps in order to facilitate movement of traps among sites. The bait type NEON uses for capture is a seed mixture of millet and sunflower seeds, as it is easy to transport, and less attractive of unwanted visitors than the typical bait, peanut butter. NEON's marking method is ear tagging. The ear tagging is conspicuous and is placed at the base of the ear. Ear tagging can be used to conduct behavioral studies and allows for the recording of the mammal's ear length when tagging. Lastly, NEON's trap array employs 10 X 10 trapping grids, with 10m spacing and 1 trap per station. This design appropriately yields for a sampling area of about 1ha. 

Using this small mammal data collected and provided by NEON this report will use R studio to visualize results and test how the distrubition of weight varies across NEON sites. To begin, the small mammal NEON data set was read into R studio. From this all the siteID's and small mammal weight's were correlated as a box plot to look at the average small mammal weights of all NEON sites. After viewing the boxplot of all the NEON site's average small mammal weights the sites with the highest average small mammal weight and the lowest average small mammal weight were then both analyzed.The site with the highest average weight was the LBJ National Grassland ("CLBJ" site) and the site with the lowest average weight was the Smithsonian Environmental Research Center ("SERC" site).These two sites were explored further using two seperate boxplots to look at why one site had a particularly high/low average weight and to ask if it was just that one site that caused the average weight to be high/low. This was done by taking the scientific names of the species at these sites against the average weight of the small mammals. Lastly, one of the "CLBJ" site's species and one of the "SERC" site's species were choosen to study further, in order to ask why the "CLBJ" site had a lot of variation in weight from species to species, and opposingly, why the "SERC" site had a much more constant weight across the board with it's varying species. The "CLBJ" species analyzed further was *Neotoma micropus* and *Mus musculus* was analyzed for the "SERC" site. One boxplot was made for the weight of *Neotoma micropus*, correlating the species present at all of the NEON sites against weight, and an additional boxplot was made for the same purpose just using the species *Mus musculus* instead.

Describe where the data came from, including how they were collected in the field (information is in the pdf file in the data folder), some information about those sites and why you chose them, and how you are analyzing them (e.g. using R to visualize the results and the statistical tests you will use).

# Results
```{r load-packages}
library("dplyr")
library("ggplot2")
library("readr")
```

Read in the small mammal data set:
```{r load-in-data}
# read in the small mammal data from raw data directory
mam_pertrapnight <- read_csv("data/raw_data/mam_pertrapnight.csv")
```

Box plot for the weight of small mammals at all NEON sites:
```{r boxplot-of-small-mammal-weight}
# weight of small mammals at all sites in the NEON data
ggplot(mam_pertrapnight, aes(x = siteID,
                             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Small Mammals at all NEON sites") +
  xlab("Site ID") +
  ylab("Weight (grams)")
```

Box plot for the weight of small  mammals at "CLBJ" NEON site:
```{r boxplot-for-CLBJ-site}
# make a boxplot for the weight of small mammals at "CLBJ" NEON site
mam_pertrapnight %>%
  filter(siteID == "CLBJ") %>%
  filter(!is.na(scientificName)) %>%
  ggplot(aes(x = scientificName,
             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Small Mammal Species at CLBJ NEON site") +
  xlab("Species") +
  ylab("Weight (grams)")
```

Boxplot for the weight of small  mammals at "SERC" NEON site:
```{r boxplot-for-SERC-site}
# make a boxplot for the weight of small mammals at "SERC" NEON site
mam_pertrapnight %>%
  filter(siteID == "SERC") %>%
  filter(!is.na(scientificName)) %>%
  ggplot(aes(x = scientificName,
             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Small Mammal Species at SERC NEON site") +
  xlab("Species") +
  ylab("Weight (grams)")
```

```{r boxplot-of-neotoma-weight-all-sites}
mam_pertrapnight %>%
  filter(scientificName == "Neotoma micropus") %>%
  ggplot(aes(x = siteID,
             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Neotoma micropus Species at all NEON Sites") +
  xlab("Site") +
  ylab("Weight (grams)")
```

```{r boxplot-of-mus-weight-all-sites}
mam_pertrapnight %>%
  filter(scientificName == "Mus musculus") %>%
  ggplot(aes(x = siteID,
             y = weight)) +
  geom_boxplot() +
  ggtitle("Weight of Mus musculus Species at all NEON Sites") +
  xlab("Site") +
  ylab("Weight (grams)")
```

Add a number of code chunks in the Results section. These should read in, subset and plot the data as needed (no need to save any figures to pdf, since they will be put into the rendered document when you click 'knit'), and, for any hypotheses that you want to test, an appropriate statistical test. You should have at least 5 figures and/or tables in this section, which you use to test your hypotheses against the dataset.

# Discussion

Discuss the results you found in the context of what you know about the site/sites and the predominant types of vegatation in those locations. Think also about the community dynamics of the mammal community at each of the sites, in the context of what you know about different types of species interactions, food webs, trophic transfer, etc.

# Sources Cited

